---
title: "P8105-HW2"
author: "Zhe Chen"
date: "2020/9/28"
output: github_document
---

## Problem 1

### Include relevent library

```{r}
library(tidyverse)
library(readxl)
```


### Import data:

```{r, message = FALSE}
trashwheel =
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = cell_cols("A:N")) %>% janitor::clean_names() %>%
drop_na(dumpster) %>%
mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )

```

### Data of Precipitation for 2018

```{r}
precip_2018 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
```

### Data of Precipitation for 2017

```{r}
precip_2017 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

### Combining 2018 and 2017 data

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip%>% filter(year == 2018) %>% pull(total) %>% sum()` inches.












