---
title: "P8105-HW2"
author: "Zhe Chen"
date: "2020/9/28"
output: github_document
---

## Problem 1

### Include relevent library

```{r}
library(tidyverse)
library(readxl)
```


### Import data

```{r, warning = FALSE}
trashwheel =
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = cell_cols("A:N")) %>% janitor::clean_names() %>%
drop_na(dumpster) %>%
mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )

```

### Data of Precipitation for 2018

```{r}
precip_2018 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
```

### Data of Precipitation for 2017

```{r}
precip_2017 = 
	read_excel(
		"./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

### Combining 2018 and 2017 data

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip%>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

### Import and Clean the Data

```{r, warning= FALSE}
# import and initial cleaning
NYC_Transit =
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% #clean the name
  select(line, name = station_name, route1:route11, station_latitude, station_longitude, entrance_type , entry, vending, ada) %>% #select meaningful variables
   mutate_at(vars(route1:route11), funs(as.character)) #convert to character for further cleaning

# make the data more tidy
NYC_Transit_tidy = 
  NYC_Transit %>%
  # merging variables to a single variable
  pivot_longer( 
      route1:route11,
      names_to = "routes",
      values_to = "routes served",
      names_prefix = "route"
    ) %>%
  drop_na("routes served") # abandon observations with missing values
NYC_Transit$entry = ifelse(NYC_Transit$entry == "YES", TRUE, FALSE) # convert entry to logical type
head(NYC_Transit_tidy) 
```

  Above is the first 10 lines from the data set some cleaning. Our data set, NYC_Transit, contains the information about entrance, exit for each subway in NYC. We cleaned the data set in several steps:
  
  * make the variable name to a uniform format;
  * keep variables that are useful(line, station name, routes, the latitude and the longitude of the station, entrance type, entry, vending, ADA), and rename some variables;
  * merge "route" variables into one variable "route" with variable "route served" which contains the transfer information and take off observations with missing values in "route served" variable; 
  * convert entry's data type to logical.
  
  After cleaning, we have a more tidy data set with dimension of `r nrow(NYC_Transit)*ncol(NYC_Transit)`. This data set is pretty tidy now.

### Reformat the data

```{r}
#
```







